---
title: "STAT5002 Lab9 Worksheet"
subtitle: "One-sample T-test"
author: "© University of Sydney STAT5002"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

Below is the corrected version of the first column of the data (labelled "Dextro-") and the the second column labelled "Laevo-" in R. We will use these two variables in this workshop. 

```{r}
dextro = c(.7, -1.6, -.2, -1.2, -.1, 3.4, 3.7, .8, 0, 2)
laevo = c(1.9, .8, 1.1, .1, -.1, 4.4, 5.5, 1.6, 4.6, 3.4)
```


# The `dextro` data

## Determine the sample mean and SD (note the original paper uses the population SD)

```{r}
# type your code here
       
```
      
## Is it reasonable to assume that this is a sample from an "approximately normal box"? Use some plots to check this.

```{r}
# type your code here
       
```

## Provide a 95% confidence interval for the average increase in hours' sleep using Dextro-.

```{r}
# type your code here
       
```

## Would we reject a two-sided test that Dextro makes no difference to the duration of sleep at the 5% level of significance? Explain.
	
# The `laevo` data

## Verify the calculations of the sample mean and SD

```{r}
# type your code here
       
```

## Is it reasonable to assume that this is a sample from an "approximately normal box"? Use some plots to check this.

```{r}
# type your code here
       
```

## Provide a 95% confidence interval for the average increase in hours' sleep using Laevo-.

```{r}
# type your code here
       
```

## Would we reject a two-side test that Laevo makes no difference to the duration of sleep at the 5% level of significance? Explain.

# Are the two treatments being equally effective? 

Are your results consistent with the two treatments being equally effective? 
Discuss with reference to **overlap between confidence intervals**.


# Test the difference

Repeat the steps of question 1 again but this time using the third column of "Laevo -- Dextro" differences for each patient.

- You can obtain a vector of differences by 

```{r}
d = laevo - dextro 
```

## Verify the calculations of the sample mean and SD of the difference

```{r}
# type your code here
       
```

## Is it reasonable to assume that these differences are like a sample from an "approximately normal box"? Use some plots to check this.

```{r}
# type your code here
       
```

## Provide a 95% confidence interval for the average difference in increase in hours' sleep:

```{r}
# type your code here
       
```

## Is this consistent with the two treatments being equally effective? In your answer, refer to your answer to question 3 above.


# Perform a formal t-test on the difference.

Perform a formal t-test that the "true" average difference over all (potential) patients is zero. Stating and checking any required assumptions. Will you reject based on a 5\% level of significance? 

## Write down the testing procedure step-by-step

- Write your code below for the P-value

```{r}
# type your code here
       
```

## Check your answers using `t.test()`.

```{r}
# type your code here
       
```


## Using simulation

Now use a bootstrap simulation to construct a 95\% confidence interval and perform a test. 
How do your results here differ from those given by `t.test()`, and why? 

- Simulate T-statistics of 10,000 samples drawn from the surrogate box (the Laevo -- Dextro difference)

```{r}
n = length(d)
box.guess = d # this is the surrogate box
sim.T.stats=0
# write your code here

```


- Plot the histogram of the simulated T-statistics and compare it with the theoretical Student's T-distribution. What is your observation? 
  - Note that `curve(dt(x, df=*), add=T, lty=2)` plots the Student's T-distribution with `*` degrees of freedoms on top of the histogram

```{r}
# write your code here

```

- Calculate P-value based on the simulated distribution.

```{r}
# write your code here

```

- Construct a 95% confidence interval, using the lower and upper 2.5% points of the simulated statistics (note that the simulated distribution is not symmetric).


```{r}
# write your code here

```


