---
title: "STAT5002 Lab7 Worksheet"
subtitle: "Confidence Intervals and Z-test"
author: "© University of Sydney STAT5002"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      # - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fas fa-chart-bar"></i> Same-Sex Marriage Survey

Between 12 September and 7 November 2017 the Australian Government conducted a national postal survey in order to gauge the Australian community's opinion on changing legislation about same-sex marriage.

In the lead-up to the postal survey, various organisations conducted their own polls to infer how the final survey might turn out. [One such poll](https://www.theguardian.com/australia-news/2017/aug/21/most-support-marriage-equality-and-80-plan-to-vote-in-survey-guardian-essential-poll/) returned the following results:

| In favour of proposed change | Against proposed change | Don't know | 
|:----------------------------:|:-----------------------:|:----------:|
|   1036                       | 581                     |   200      |

What can we infer from this about the final results of the official postal survey?

## Q1.3: Confidence interval {-}

- **Determine a Wilson's 99% confidence interval for $p$ based on the observations in the previous part. Hint:** use `binom.confint()`.

```{r}
require(binom)    
# write code here

```

- **Perform a "sanity check" that the endpoints of the interval are such that the observed value $\bar x$ is "right on the edge" of a corresponding 99% prediction interval.**

  - You need to identify the appropriate multiplier for a 99% prediction interval
  - Note that the endpoints of the Wilson interval are available via the objects `CI$lower` and
`CI$upper`:

```{r}
# write code here for the confidence interval

```

## Q1.4: Hypothesis test {-}

**The observed proportion is greater than 0.5, but is it significantly greater? Answer this question with an appropriate formal hypothesis test.**

```{r}
# write code here for the P-value


```

